Load Libraries

```{r}
library(tidyverse)
library(here)

options(scipen = 999)
```
Create theme

```{r}

visit_theme <- theme(axis.text = element_text(size = 10),
                    plot.title = element_text(size = 14),
                    panel.grid.minor.y = element_blank())

visit_colour <- c("#650360", "#005EB8")
#other colour to consider #211511

```

Load data
```{r}
activities <- read_csv(here("clean data/activities.csv"))

locations <- read_csv(here("clean data/location.csv"))

travel <- read_csv(here("clean data/travel.csv"))

accomodation <- read_csv(here("clean data/accomodation.csv"))

demographics <- read_csv(here("clean data/demographics.csv"))
```


# What kind of activity generates the highest income for Scotland?

```{r}

activities <- activities %>% 
  mutate(gdp_spent_millions = gbp_spent/1000000,
         visits_millions = visits/1000000)

activities %>% 
  filter(tourism_activity != "All") %>% 
  group_by(tourism_activity) %>% 
  summarise(total_income_millions = sum(gdp_spent_millions)) %>% 
  arrange(desc(total_income_millions))

activities %>% 
  filter(tourism_activity != "All") %>% 
  group_by(tourism_activity) %>% 
  summarise(total_income_millions = sum(gdp_spent_millions)) %>% 
  arrange(desc(total_income_millions)) %>%
  mutate(tourism_order = 1:n()) %>%
  mutate(tourism_order = as.character(tourism_order)) %>% 
  ggplot(aes(x = total_income_millions, y = fct_reorder( tourism_activity, total_income_millions), fill = forcats::fct_inorder( tourism_order))) +
  geom_col()+
  scale_fill_manual(values = rep(visit_colour, length.out = 15))+
  theme(legend.position = "none")+
  labs(x = "Total spend (£millions)",
       y = "Type of tourism activity",
       title = "Total spent on activity")+
  visit_theme

```
The 3 activities that generate the most income for Scotland are eating out, shopping for items that would not generally be bought and visits with friends and family all with over £10 billion generated each. This equates to over £1 billion per annum for each of these types of tourism income.

Following these nights out, outdoor leisure and general days out all generated over £5 billion between 2013 and 2019.

# What kind of activity generates is the cause of the most visits to sites in Scotland?

```{r}

activities %>% 
  filter(tourism_activity != "All") %>% 
  group_by(tourism_activity) %>% 
  summarise(total_visits_millions = sum(visits_millions)) %>% 
  arrange(desc(total_visits_millions))

activities %>% 
  filter(tourism_activity != "All") %>% 
  group_by(tourism_activity) %>% 
  summarise(total_visits_millions = sum(visits_millions)) %>% 
  arrange(desc(total_visits_millions)) %>% 
  mutate(tourism_order = 1:n()) %>%
  mutate(tourism_order = as.character(tourism_order)) %>% 
  ggplot(aes(x = total_visits_millions, y = fct_reorder(tourism_activity, total_visits_millions), fill = forcats::fct_inorder(tourism_order))) +
  geom_col()+
  scale_fill_manual(values = rep(visit_colour, length.out = 15))+
  theme(legend.position = "none")+
  labs(x = "Total visits (millions)",
       y = "Type of tourism activity",
       title = "Total visits to activity")+
  visit_theme

```
The largest reason for visits to sites in Scotland is visiting friends and family with over 300 million visits.

Following this going out for a meal and outdoor leisure activities were both identified as the reason for over 150 million visits.

General days out, nights out and shopping also accounted for more than 100 million visits each.

# What kind of activity generates the most spend per visit?

```{r}

activities %>% 
  mutate(spend_per_visit = gdp_spent_millions / visits_millions) %>% 
  filter(tourism_activity != "All") %>% 
  group_by(tourism_activity) %>% 
  summarise(average_spend_per_visit = mean(spend_per_visit)) %>% 
  arrange(desc(average_spend_per_visit))

activities %>% 
  mutate(spend_per_visit = gdp_spent_millions / visits_millions) %>% 
  filter(tourism_activity != "All") %>% 
  group_by(tourism_activity) %>% 
  summarise(average_spend_per_visit = median(spend_per_visit)) %>% 
  arrange(desc(average_spend_per_visit)) %>% 
  mutate(tourism_order = 1:n()) %>%
  mutate(tourism_order = as.character(tourism_order)) %>% 
  ggplot(aes(x = average_spend_per_visit, y = fct_reorder(tourism_activity, average_spend_per_visit), fill = forcats::fct_inorder(tourism_order))) +
  geom_col()+
  scale_fill_manual(values = rep(visit_colour, length.out = 15))+
  theme(legend.position = "none")+
  labs(x = "Average spend per visit",
       y = "Type of tourism activity",
       title = "Average spend by activity")+
  visit_theme  

```

When looking at the data by the amount spent per visit we can see that shopping for items that you do not regularly buy is highest. The days out to beauty/health centre/spas, meals and nights out all generated an average of over £50 per visit. 

Based on this analysis we should look to increase the number of trips people take to these type of tourism activities to help generate additional income throughout Scotland.

# What key demographics shoudl be focused on? Who spends the most and who visits the most?



# What type of locations receive the most visits and where do people spend the most?

```{r}

locations %>% 
  filter(type_of_location_visited != "All areas") %>% 
  mutate(visits_millions = visits / 1000000) %>% 
  group_by(type_of_location_visited) %>% 
  summarise(total_visits_millions = sum(visits_millions)) %>% 
  ggplot(aes(x = total_visits_millions, y = fct_reorder(type_of_location_visited, total_visits_millions))) +
  geom_col()

locations %>% 
  filter(type_of_location_visited != "All areas") %>% 
  mutate(spend_millions = gbp_spent / 1000000) %>% 
  group_by(type_of_location_visited) %>% 
  summarise(total_spent_millions = sum(spend_millions)) %>% 
  ggplot(aes(x = total_spent_millions, y = fct_reorder(type_of_location_visited, total_spent_millions))) +
  geom_col()

locations %>% 
  mutate(spend_per_visit = gbp_spent / visits) %>% 
  filter(type_of_location_visited != "All areas") %>% 
  group_by(type_of_location_visited) %>% 
  summarise(average_spend = median(spend_per_visit)) %>% 
  ggplot(aes(x = average_spend, y = fct_reorder(type_of_location_visited, average_spend))) +
  geom_col()

```
Cities and large towns are the most visited locations in Scotland. The number of visits approximately lines up with the expected size/populations of the locations with some possible exceptions around seaside towns.  With visits friends and family being the top reason for tourism activities it makes sense that visits to the most populated areas would be most common.

By a considerable margin the most money is spent in cities and large towns. This is partially due to the fact that this is where most visits occur alongside the fact things are often more expensive in Cities leading to a higher spend per visit. All other locations average a spend of between £25 and £40 per visit.


# Is there a particular method of travel visitors arrive by? 

```{r}
travel %>% 
  filter(method_of_travel != "All") %>% 
  ggplot(aes(x = visits, y = fct_reorder(method_of_travel, visits))) +
  geom_col()

travel %>% 
  mutate(car_or_not = case_when(method_of_travel == "Car - hired" ~ "Car",
                                method_of_travel == "Car - own/friends/family" ~ "Car",
                                TRUE ~ "Not Car")) %>% 
  filter(method_of_travel != "All") %>% 
  ggplot(aes(x = visits, y = fct_reorder(car_or_not, visits))) +
  geom_col()

travel %>% 
  mutate(public_or_not = case_when(method_of_travel == "Boat/Ship/Ferry" ~ "Public",
                                   method_of_travel == "Bus or Coach - Minibus" ~ "Public",
                                   method_of_travel == "Bus or Coach - Organised Tour" ~ "Public",
                                   method_of_travel == "Bus or Coach - Regular" ~ "Public",
                                   method_of_travel == "Plane" ~ "Public",
                                   method_of_travel == "Taxi" ~ "Public",
                                   method_of_travel == "Train" ~ "Public",
                                   method_of_travel == "Tram" ~ "Public",
                                   method_of_travel == "Tube" ~ "Public",
                                   TRUE ~ "Not Public"
                                   )) %>% 
  filter(method_of_travel != "All") %>%
  filter(public_or_not == "Public") %>% 
  ggplot(aes(x = visits, y = fct_reorder(method_of_travel, visits))) +
  geom_col()

```

Car is far and away the most popular form of transport to arrive at these locations by. In fact there is almost double the amount of visits made by car than all other methods of transport combined.

Within those people who traveled by public transport train and bus/coach (regular) were the most common transportation taken for visits to tourism activities. 




# What type of accomodation is most popular?

```{r}

accomodation %>% 
  filter(location == "All" & weekday_weekend == "All", size_of_accommodation == "All") %>% 
  group_by(accommodation_type_and_occupancy, year) %>% 
  summarise(average_occupancy = median(percent_occupancy)) %>% 
  ggplot(aes(x = year, y= average_occupancy, colour = accommodation_type_and_occupancy))+
  geom_line()

```

Due to the impact of covid-19 this data has been severely impacted with very few stays in any kind of accommodation in 2020. While 2021 shows some level of recovery the majority of accommodation types do not appear to have fully recovered. The data used to answer any business questions will therefore be cut off at 2019 before these unprecedented disruptions took place.

The type of accommodation which shows the least impact to covid-19 appears to be caravan and camping parks. With these being full outdoors this type of accommodation will have seen less restrictions and possibly have been more able to stay open and seemed more attractive to prospective customers.


```{r}

accomodation_to_2019 <-  accomodation %>% 
  filter(year != 2021, year != 2022)

accomodation %>% 
  filter(location == "All" & weekday_weekend == "All", size_of_accommodation == "All") %>% 
  group_by(accommodation_type_and_occupancy) %>% 
  summarise(average_occupancy = median(percent_occupancy)) %>% 
  ggplot(aes(x = average_occupancy, y = accommodation_type_and_occupancy))+
  geom_col()

```
The data is not available for the total number of stays in the various types of accommodation so occupancy rates will be used to judge popularity, with more occupied accommodation being deemed more popular than less occupied.

Over the full dataset across all sizes of accommodation and location bed occupancy is highest in hotels indicating these are the most popular place to stay.

With additional data it would be very beneficial to be able to see areas with high occupancy rates which could be somewhere which could sustain additional hotels to increase overall visits. It would also be of interest to see how the occupancy runs throughout the year in various types of accommodation. It would be expected that accommodation like camping and caravan parks would see a sharp decline over the winter period while this would not be as extreme for hotels.

# Is there a difference is which accomodation is popular at differenet locations?


```{r}
accomodation

accomodation %>% 
  filter(location != "All") %>% 
  filter(weekday_weekend == "All", size_of_accommodation == "All") %>% 
  group_by(accommodation_type_and_occupancy, location) %>% 
  summarise(average_occupancy = mean(percent_occupancy)) %>% 
  ggplot(aes(x = average_occupancy, y = accommodation_type_and_occupancy))+ 
  geom_col()+
  facet_wrap(~location)


```

# Is there a particular demographic of visitors that we should be focusing on? Who spends the most and who visits the most?

```{r}

demographics

demographics %>% 
  filter(category == "social_grade") %>% 
  ggplot(aes(x = year, y = gbp_spent))+
  geom_line()+
  facet_wrap(~type)

```


# What insights can we gain from regional tourism? Where are our visitors from and how does the money spent/ nights stayed and visits made compare?

# How have general tourism rates changed over time?

# Is there any way to predict spending and visits for Visit Scotland?
